%%% WETLAB STUFF
\paragraph{Strains and Media.}
We used the bioluminescent strain \textit{Escherichia coli} MG1655 \(\Delta\)galK::(kan\(^R\)-luxCDABE) constructed previously \cite{Muetter2026}.
Cultures were grown in LB medium (Sigma L3022).

\paragraph{Drug preparation.}
We explored drug interactions among six drugs (\autoref{tab:drugs}), resulting in \num{15} drug--drug pairs.
For each pair, one compound was designated drug~A and the other drug~B.
Each compound was prepared as a \(\times 20\) stock of its highest working concentration.
We then performed a twofold dilution series in a 12-column deepwell plate, resulting in \num{11} diluted concentrations and one drug-free column.
For drug~A, we transferred \SI{125}{\micro\liter} per well into a 96-well plate (subreservoir~A) with a horizontal concentration gradient.
For drug~B, we transferred \SI{50}{\micro\liter} per well into two 96-well plates (antibiotic reservoir plates I and II) with vertical concentration gradients (each \(6\times 12\) layout).
All plates were stored at \SI{-80}{\celsius} to minimise degradation over time, accepting a one-time freeze--thaw-related degradation.

%% DRUG TABLE
\begin{table}
  \caption{Drugs used in the experiments, the solvent, the estimated zMIC (by fitting the single-drug pharmacodynamic curves), and the supplier with catalogue number.}
  \centering
  \small
  \input{tables/drug_table}
  \label{tab:drugs}
\end{table}

\paragraph{Dose response assays.}
For each assay, we prepared four overnight cultures grown for \SI{14}{\hour}.
Cultures were distributed to fill an entire 384-well source plate (Greiner 781073), so that the replicates formed a \(2\times2\) block format (e.g. C1--rep1, C2--rep2, D1--rep3, D2--rep4).
We then prefilled two white 384‑well assay plates (Greiner 781073) with \SI{54}{\micro\liter} LB per well and transferred inocula from the source plate to both assay plates (I, II) using an Evo 200 liquid handling platform (Tecan) with a pintool.
The dilution is approximately \(1{:}150\), although we observed substantial well-to-well variation in the effective inoculum size (see below).
Plates were incubated for \SI{2}{\hour} to reach exponential phase.
Simultaneously, subreservoir~A and both antibiotic‑reservoir plates were thawed.
\SI{50}{\micro\liter} from each well of the subreservoir~A were transferred to the corresponding wells of both antibiotic‑reservoir plates, generating a \num{10}\,x mixture of drugs A and B.
For both assay plates, the assay started as follows: a baseline luminescence reading was taken, \SI{6}{\micro\liter} of the \num{10}\,x drug mixture was added (defining \(t=0\) at dosing for the respective wells), and a second luminescence reading was taken.
Subsequently, we alternated between reading assay plates I and II for a total duration of \SI{5}{\hour}.
Luminescence was recorded with an Infinite~F200 plate reader (Tecan) using a \SI{250}{\milli\second} integration time.

\paragraph{Data preprocessing.}
To estimate stray.light contamination, we conducted a calibration experiment in which six source wells (E5, E12, E20, L5, L12, L20) contained stationary-phase cultures while all other wells remained empty.
From a single luminescence read of the full plate, we constructed a distance-dependent stray-light kernel (\autoref{fig:light_distribution}) and corrected each well by subtracting the summed contributions from neighbouring wells (restricted to distances $d\le 3$, \autoref{eq:noise_corr}).
We defined the lower limit of detection as \(I_{\mathrm{LoD}}=10\,\mathrm{RLU}\) and the upper limit of detection as \(10^{6}\,\mathrm{RLU}\).
We defined the common analysis horizon \(T\) as the earliest time point at which any untreated control well exceeded the upper detection limit, yielding \(T\approx\SI{2}{\hour}\).
For each well, we computed the fraction of observations up to \(T\) below \(I_{\mathrm{LoD}}\).
Wells with more than $20\%$ of observations below \(I_{\mathrm{LoD}}\) were excluded from all subsequent analyses.
For retained wells, values below \(I_{\mathrm{LoD}}\) were censored by replacing them with \(I_{\mathrm{LoD}}\).
For numerical integration, we interpolated the log-normalised signal to obtain values on the shared horizon \(t=T\) (see also \autoref{ssec:lightnoise}).

\paragraph{Time-weighted net growth rate.}
We define \(\psi\) as the linearly time-weighted net growth rate (SI: \autoref{eq:psi_weighted}) over the common time horizon \(T\).
We infer \(\psi\) by calculating the scaled area under the log-normalised light intensity \(Y(t)=\ln\!\big(I(t)/I(0)\big)\) (SI: \autoref{eq:norm}, \autoref{eq:Ydef}),
\begin{equation}
  \psi
  =
  \frac{2}{T^{2}} \int_{0}^{T}Y(t)\,\mathrm{d}t.
  \label{eq:effective_psi}
\end{equation}

\paragraph{Classification of inhibitory and sub-inhibitory conditions.}
We classify each checkerboard condition as inhibitory, sub-inhibitory, or not significantly different from zero by testing whether replicate time-weighted growth-rate estimates \(\psi(c_A,c_B)\) are consistently positive or negative.
With four replicates, the binomial probability of observing all positive (or all negative) signs with equal probability is \(0.5^4=0.0625\).
This classification is used to annotate the \(\psi\) heatmaps (\autoref{fig:psi_chl_tet}, \autoref{fig:psi}) and to define inhibitory and sub-inhibitory regime pools.

\paragraph{Single-drug pharmacodynamic curve fitting.}
For each drug, we aggregated all single-drug data across experiments and fitted pharmacodynamic curves \(f(c)\) to the corresponding time-weighted growth-rate estimates \(\psi\):
\begin{equation}
  f(c) = \psi_{0} - \frac{(\psi_{0}-\psi_{\min})(c/\mathrm{zMIC})^{\kappa}} {(c/\mathrm{zMIC})^{\kappa}-(\psi_{\min}/\psi_{0})}, \quad [0,\infty)\rightarrow[\psi_{\min},\psi_{0}]
  \label{eq:single_drug_pd}
\end{equation}
Here, \(\psi_{0}\) denotes the maximum and \(\psi_{\min}\) the minimum time-weighted growth rate, corresponding to \(c=0\) and \(c\to\infty\), respectively.
\(\kappa\) denotes the Hill coefficient and \(\mathrm{zMIC}\) the concentration at which \(f(\mathrm{zMIC})=0\).
The model parameters $psi_{0}$, $\psi_{\min}$, $\kappa$, $\mathrm{zMIC}$ were jointly estimated by least-squares fitting of \autoref{eq:single_drug_pd}.

\paragraph{Distribution of treatment effects.}
We define the treatment effect $\tau$ as the linearly time-weighted average difference between the untreated and treated instantaneous net growth rates over the common time horizon \(T\) (SI: \autoref{eq:intYdiff}).
Equivalently, \(\tau\) is the scaled area between the control and treatment trajectories on the log-normalised scale (illustrated in \autoref{fig:peptide-interactions}b).
Because \(\tau\) is a comparative measure between one untreated and one treated trajectory, we bootstrap treated--control well pairs with replacement (\(B=200\); details in \autoref{ssec:bootstrap_interactions}).
For each condition \((c_A,c_B)\) and draw \(b\), we compute \(\tau_b(c_A,c_B)=\psi_b(\emptyset)-\psi_b(c_A,c_B)\), yielding the distribution of treatment effects $\symbf{\tau}(c_A, c_B)$.
Throughout this manuscript, distributions are denoted by bold symbols.
We classify such bootstrap distributions as significant if the central \(95\%\) bootstrap interval excludes zero, and as not significant otherwise.

\paragraph{Bliss-based interaction score \(\mu\).}
We define the Bliss based interaction score as the normalised deviation from the Bliss-prediction for combined treatment effects (see \autoref{ssec:bliss} and \autoref{eq:tau_bliss_main}):
\begin{equation}
  \mu_b(c_A,c_B)  =
  \frac{\tau_b(c_A)+\tau_b(c_B)-\tau_b(c_A,c_B)}
  {\mathrm{median}\!\bigl(\symbf{\tau}(c_A)\bigr)+\mathrm{median}\!\bigl(\symbf{\tau}(c_B)\bigr)}.
  \label{eq:mu}
\end{equation}
To avoid misinterpreting noise as interaction, we only evaluate \(\nu(c_A,c_B)\) if \(\symbf{\tau}(c_A,c_B)\) is significantly greater than zero and at least one of \(\symbf{\tau}(c_A)\) or \(\symbf{\tau}(c_B)\) is significantly greater than zero.

\paragraph{Loewe-based interaction score \(\nu\).}
We define a Loewe interaction score (details in \autoref{ssec:loewe}) that quantifies deviations from dose equivalence as
\begin{equation}
  \nu_b(c_A,c_B)
  =
  \frac{c_A}{f_A^{-1}\!\big(\psi_b(c_A,c_B)\big)}
  +
  \frac{c_B}{f_B^{-1}\!\big(\psi_b(c_A,c_B)\big)}
  -1.
  \label{eq:fici}
\end{equation}
Here \(f_i\) denotes a pharmacodynamic function mapping concentration to net growth, \(f_i(c)=\psi\).
The inverse \(f_i^{-1}(\psi)\) is the corresponding single-drug equivalent concentration of drug \(i\).
Our definition of \(\nu\) is analogous to the combination index \((\mathrm{CI}-1)\) \cite{Chou1984}, but uses a different pharmacodynamic model \(f_i\) (either \autoref{eq:psi_inverse_closed_form} or inferred numerically from the single-drug edges of the surface splines).
To avoid misinterpreting noise as interaction, we only evaluate \(\nu(c_A,c_B)\) if \(\symbf{\tau}(c_A,c_B)\) is significantly greater than zero and at least one of \(\symbf{\tau}(c_A)\) or \(\symbf{\tau}(c_B)\) is significantly greater than zero.

\paragraph{Polar reparametrization.}
We normalise concentrations of drug \(i\) using the corresponding \(\mathrm{zMIC}_i\) estimates obtained from pharmacodynamic curve fits (\autoref{tab:drugs}): \(z_i=c_i/\mathrm{zMIC}_i\).
We then define polar coordinates based on the normalised concentrations:
\begin{equation}
  z=\sqrt{z_{A}^{2}+z_{B}^{2}},
  \qquad
  \phi=\arctan2\!\bigl(z_{B},z_{A}\bigr),
  \label{eq:polar_coordinates}
\end{equation}
with \(\phi\) corresponding to the \emph{mixing angle} and \(z\) to the \emph{combined dose}.
These coordinates were used to plot polar pharmacodynamic curves, which show the treatment effect over the combined dose \(z\) for a fixed mixing angle \(\phi\).
These curves represent cross-sections through the surface splines at fixed \(\phi\).

\paragraph{Inoculum effects.}
We noticed a much larger-than-expected variation in pre-treatment light intensity \(I_0\) in our data, which we use as a proxy for inoculum size.
To assess the impact of this variation on the results, we regressed \(\psi\) for each single-drug and concentration on the pre-treatment signal \(I_0\) (\autoref{ssec:inoculum}).
We found negligible inoculum effects for AMO, CHL, PEN, and TET, but substantial effects for COL and POL at intermediate concentrations (see \autoref{fig:inoculum}).
Because the size of the inoculum did not show a significant directional trend along the concentration index (\(p=0.092\)), the variance of the inoculum mainly adds noise at intermediate concentrations of COL and POL, contributing to an increased scatter in \autoref{fig:single-pdcurves}c,e.
