% !TEX root = ../main.tex

\documentclass[../main.tex]{subfiles}
\input{commands.tex}

\title{Antimicrobial Combination Effects at Sub-inhibitory Doses do not Reliably Predict Effects at Inhibitory Concentrations}

\manuscriptauthors{Malte Muetter\textsuperscript{a}, Daniel C.\ Angst\textsuperscript{a}, Roland R.\ Regoes\textsuperscript{a}, Sebastian Bonhoeffer\textsuperscript{a}}

\affil{a}{Institute of Integrative Biology, Department for Environmental System Science, ETH Zurich, 8092 Zurich, Switzerland}

\manuscriptpublished{This manuscript was published as a preprint on \textit{bioRxiv} on February 8, 2026 (doi: 10.64898/2026.02.07.703730) and is currently under review at Molecular Systems Biology.}

\authorcontributions{%
  MM, DCA, RRR, and SB designed research; MM performed experiments, developed the mathematical model, and analyzed data; MM, DCA, RRR, and SB wrote the manuscript.
}

\manuscriptabstract{
  Assessing whether drug combinations synergise or antagonise is difficult for several reasons: (i) measuring bacterial death rates at clinically relevant drug concentrations is methodologically challenging, (ii) there is no unifying definition of what constitutes synergistic or antagonistic interactions, and (iii) both synergism and antagonism may be concentration and mixing ratio dependent.
  To assess how well sub-inhibitory measurements predict inhibitory behaviour, we quantified drug interactions for \num{15} pairwise drug combinations on a concentration checkerboard covering a wide range of inhibitory and sub-inhibitory concentrations.
  To this end, we recorded \num{8640} time-resolved luminescence trajectories that have been shown to track bacterial population declines.
  To handle time-varying treatment effects and allow fair comparisons between drugs with distinct killing dynamics, we introduced a growth-integrated rate-like metric \(\psi\) to summarise each trajectory, and assigned interaction types (synergistic/independent/antagonistic) based on Bliss independence and Loewe additivity.
  We found that the interaction type frequently changes as the concentration increases from sub-inhibitory to inhibitory concentrations.
  Moreover, interaction type depended on the mixing ratio, implying that single-point sub-inhibitory measurements are not sufficient to predict interactions at clinically relevant concentrations.
}

\begin{document}
\manuscriptfront

\section{Introduction}
\input{chapter_4/11_introduction.tex}

\section{Results}
\input{chapter_4/12_results.tex}

\section{Discussion}
\input{chapter_4/13_discussion.tex}

\section{Methods}
\input{chapter_4/14_methods.tex}

\section*{Data, Materials, and Software Availability}
Experimental datasets and code are available at Zenodo (DOI: \href{https://doi.org/10.5281/zenodo.18374151}{10.5281/zenodo.18374151}).

\section*{Acknowledgements}
We thank ETH Zurich for funding this work.
During manuscript preparation, we used OpenAI’s ChatGPT for editorial assistance (coding, grammar, and proofreading).
\end{document}
