\chapter*{Acknowledgements}
First and foremost, I would like to thank Daniel Angst, Sebastian Bonhoeffer, and Roland Regoes for being amazing supervisors.
Daniel, thank you for answering hundreds of questions without ever losing patience, for teaching me everything I needed to know in the wet lab, and for being such a fun and wonderful person.
Sebastian and Daniel, thank you for allowing me to join the group despite my initial ignorance, and for giving me the time, support, and environment I needed to develop.
Sebastian and Roland, thank you for always having an open door despite being incredibly busy and for the many interesting discussions and insightful advice over the years.
Your support and guidance have been invaluable to me, and I am truly grateful for the opportunity to work with you.
I also thank Tobias Bollenbach for being part of my committee and for the pleasant and always encouraging correspondence.

Sebastian and Roland, thank you for creating the Theoretical Biology Group by bringing people from all disciplines together to create something new.
The group is truly amazing, and I want to thank every member for the discussions, feedback, and for making everyday life both stimulating and fun.
It has been a privilege to be part of this group, and I enjoyed all the lunches and coffee breaks with you.

Finally, I am deeply grateful to my family.
Thank you to my parents for their constant encouragement and unwavering support over the years, for always being a place of trust and stability, and for believing in me.
Most importantly, I thank my wife Hanna for her love, patience, and strength.
Your support made this journey possible in more ways than I can express.
Thank you, Louis, for bringing so much love and joy into my life, just by being you.
