\documentclass[../main.tex]{subfiles}


\newcommand{\siref}[1]{\textcolor{blue}{S#1}}
\newcommand{\sifig}[1]{\textcolor{blue}{Figure~S#1}}
\newcommand{\sitab}[1]{\textcolor{blue}{Table~S#1}}
\newcommand{\sieq}[1]{\textcolor{blue}{Equation~S#1}}
\newcommand{\sisec}[1]{\textcolor{blue}{#1}}
\newcommand{\panelref}[2]{\autoref{#1}\textcolor{blue}{#2}}


\begin{document}

% Paper-specific metadata
\chapter{The Impact of Treatment Strategies on the Epidemiological Dynamics of Plasmid-Conferred Antibiotic Resistance}

% Optional: include abstract if needed in thesis
\begin{abstract}
The issue of antibiotic resistance is a critical concern for public health, prompting numerous investigations into the impact of treatment strategies on preventing or slowing down the emergence of resistance.
While existing studies have predominantly focused on chromosomal resistance mutations, the consequences of often clinically more relevant plasmid-conferred resistance remain insufficiently explored.

To address this gap, we conducted three extensive \textit{in vitro} experiments utilising a liquid-handling platform.
These experiments evaluated the efficacy of five distinct treatment strategies using two antibiotics (tetracycline and ceftazidime) along with two horizontally transmissible clinical resistance plasmids conferring the respective resistances.

Among the experimentally investigated treatment strategies, combination therapy proved to be the most effective in preventing the emergence of double resistance while minimising the number of infections.

To verify the reliability of these findings, we constructed a computational model of our experiments that we parameterised using the experimental data.
We employed this model to augment the experimental data by conducting an \textit{in silico} parameter sensitivity analysis.
The sensitivity analysis corroborated our experimental results, demonstrating that combination therapy consistently outperformed other treatment strategies across a range of parameter values.
\end{abstract}

% Include the actual content from text.tex

\input{main_1/introduction.tex}
\input{main_1/results.tex}
\input{main_1/discussion.tex}
\input{main_1/methods.tex}


% Optional: acknowledgments as section
\section*{Acknowledgments}
\input{main_1/acknow.tex}

% Bibliography
\bibliographystyle{unsrtnat}
\bibliography{main_1/TSPlasmid.bib}

\end{document}
